---
title: "Reproducible_Research_Assingment_1"
author: "Paul M."
date: '2022-10-31'
output:
  pdf_document: default
  html_document: default
---


## Loading and preprocessing the data

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```

```{r, results='markup', warning=TRUE, message=TRUE}
data <- read.csv('~/activity.csv')
```


##What is mean total number of steps taken per day?
#For this part of the assignment, you can ignore the missing values in the dataset.

```{r}
daysteps <- tapply(data$steps, data$date, sum, na.rm=TRUE)
summary(daysteps)
```

#### 1. Make a histogram of the total number of steps taken each day
```{r fig.align="center"}
qplot(daysteps, xlab='Steps Per Day', ylab='Frequency', binwidth=500)

```

#### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
daystepsMean <- mean(daysteps)
daystepsMedian <- median(daysteps)

```

* Mean: `r daystepsMean`
* Median:  `r daystepsMedian`

## What is the average daily activity pattern?
```{r}
averagep<- aggregate(x=list(meanSteps=data$steps), by=list(interval=data$interval), FUN=mean, na.rm=TRUE)
summary(averagep)
```

#### 1. Make a time series plot
```{r fig.align="center"}
ggplot(data=averagep, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5 minute interval") +
    ylab("average steps") 
```

#### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r fig.align="center"}
maxsteps <- which.max(averagep$meanSteps)
timemax <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averagep[maxsteps,'interval'])
```

* Most Steps at: `r timemax`

----

## Imputing missing values
#### 1. Calculate and report the total number of missing values in the dataset 
```{r}
null <- length(which(is.na(data$steps)))
```

* Number of missing values: `r null`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
DataImputed <- data
DataImputed$steps <- impute(data$steps, fun=mean)
```


##### 4. Make a histogram of the total number of steps taken each day 
```{r fig.align="center"}
stepsImputed <- tapply(DataImputed$steps, DataImputed$date, sum)
qplot(stepsImputed, xlab='steps per day (Imputed)', ylab='Frequency', binwidth=500)
```

##### Calculate and report the mean and median total number of steps taken per day. 
```{r}
MeanImputed <- mean(stepsImputed)
MedianImputed <- median(stepsImputed)
```
* Mean (Imputed): `r MeanImputed`
* Median (Imputed):  `r MedianImputed`


----

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
DataImputed$dateType <-  ifelse(as.POSIXlt(DataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### 2. Make a panel plot containing a time series plot

```{r fig.align="center"}

avgDataImputed <- aggregate(steps ~ interval + dateType, data=DataImputed, mean)
ggplot(avgDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5 minute interval") + 
    ylab("avarage steps")
```


